---
title: "Lecture 07 - Exercise solutions"
format:
  html:
    embed-resources: true
execute:
  echo: true
  warning: true
engine: knitr
---

```{r}
library(dplyr)
```


## Exercise 1

```{r}
penguins = palmerpenguins::penguins |> as.data.frame()
```

### Select the rows for Gentoo penguins with a body mass above 5000 g, keeping only the species, island, and body_mass_g columns.

```{r}
penguins[
  penguins$species == "Gentoo" & penguins$body_mass_g > 5000,
  c("species", "island", "body_mass_g")
] |>
  str()
```

```{r}
penguins |>
  filter(species == "Gentoo", body_mass_g > 5000) |>
  select(species, island, body_mass_g)
```

### Add a column body_mass_kg containing the body mass in kilograms.

```{r}
penguins$body_mass_kg = penguins$body_mass_g / 1000
penguins |> head() 
```

```{r}
penguins |>
  mutate(
    body_mass_kg = body_mass_g
  )
```


### Calculate the mean bill length (in mm) of the Adelie penguins, ignoring missing values.

```{r}
penguins$bill_length_mm[penguins$species == "Adelie"] |>
  mean(na.rm = TRUE)
```

```{r}
penguins |>
  filter(species == "Adelie") |>
  summarize(
    species = unique(species), 
    mean = mean(bill_length_mm, na.rm=TRUE)
  )
```

## Examples

```{r}
library(dplyr)
library(nycflights13)
```

1. How many flights to Los Angeles (LAX) did each of the legacy carriers (AA, UA, DL or US) have in May from JFK, and what was their average duration?

   ```{r}
   flights |>
    filter(dest == "LAX", carrier %in% c("AA","UA","DL","US")) |>
    filter(month == 5 & origin == "JFK") |>
    group_by(carrier) |>
    summarize(
      n = n(),
      avg_duration = mean(air_time, na.rm=TRUE)
    )
   ```

2. Which plane (check the tail numbers) flew out of each New York airport the most?

   ```{r}
   flights |>
    filter(!is.na(tailnum)) |>
    count(origin, tailnum) |>
    group_by(origin) |>
    slice_max(n, n=1)
    #slice_max(n, n=1, by=origin)
   ```


3. Which 5 days should you consider flying on if you want to have the lowest possible average departure delay?

   ```{r}
   flights |>
    summarize(
      avg_dep_delay = mean(dep_delay, na.rm = TRUE),
      .by = c(month, day)
    ) |>
      slice_min(avg_dep_delay, n=5)
   ```

   ```{r}
   flights |>
    group_by(month, day) |>
    summarize(
      avg_dep_delay = mean(dep_delay, na.rm = TRUE),
      .groups = "drop"
    ) |>
      #ungroup() |>
      slice_min(avg_dep_delay, n=5)
   ```

4. Which flight has the largest arrival delay as a percentage of its scheduled air time?

   ```{r}
   flights |>
    mutate(
      arr_delay_pct = 100 * arr_delay / air_time 
    ) |>
      slice_max(arr_delay_pct, n=1) |>
      select(year:day, carrier, flight, origin, dest, arr_delay, air_time, arr_delay_pct)
   ```
