---
title: "Lecture 12 - Exercise solutions"
format:
  html:
    embed-resources: true
execute:
  echo: true
  warning: true
engine: knitr
---
```{r setup}
#| include: false
library(stringr)
```

```{python py_setup}
#| include: false
import re
```


# Exercise 1

For the vector of randomly generated names, write a regular expression that detects if the first name, the last name, either, or neither start with a vowel.

```{r}
x = c("Jeremy Cruz", "Nathaniel Le", "Jasmine Chu", "Bradley Calderon Raygoza",
      "Quinten Weller", "Katelien Kanamu-Hauanio", "Zuhriyaa al-Amen",
      "Travale York", "Alexis Ahmed", "David Alcocer", "Jairo Martinez",
      "Dwone Gallegos", "Amanda Sherwood", "Hadiyya el-Eid", "Shaimaaa al-Can",
      "Sarah Love", "Shelby Villano", "Sundus al-Hashmi", "Dyani Loving",
      "Shanelle Douglas")
```

```{python}
x = r.x
```

1. First name starts with a vowel.

   ```{r}
   str_subset(x, "^[AEIOU]")
   ```

2. Last name starts with a vowel.

   ```{r}
   str_subset(x, " [AEIOUaeiou]")
   ```


3. Either the first or last name starts with a vowel.

   ```{r}
   str_subset(x, "^[AEIOU]| [AEIOUaeiou]")
   ```


4. Neither the first nor last name starts with a vowel.

   ```{r}
   x[ !str_detect(x, "^[AEIOU]") & 
      !str_detect(x, " [AEIOUaeiou]") ]


   setdiff(x, str_subset(x, "^[AEIOU]| [AEIOUaeiou]"))

   str_subset(x, "^[AEIOU]| [AEIOUaeiou]", negate = TRUE)
   ```


# Exercise 2

Write a regular expression that extracts all of the phone numbers in `text`, then use groups to extract the area code separately from the rest of the number.

```{r}
text = c(
  "apple",
  "219 733 8965",
  "329-293-8753",
  "Work: (579) 499-7527; Home: (543) 355 3679"
)
```

```{python}
text = r.text
```

1. Extract all phone numbers.

   ```{r}
   str_extract_all(
      text,
      "\\(?\\d{3}\\)?[ -]\\d{3}[ -]\\d{4}"
   )
   ```

   
2. Area code separately from the rest of the number.

   ```{r}
   str_match_all(
      text,
      "\\(?(\\d{3})\\)?[ -]((\\d{3})[ -](\\d{4}))"
   )
   ```
